---
title: "P8105_hw2_hx2306"
author: "HuijunXiao"
date: "10/1/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
```

## Problem 1

* Step 1: Read and clean the Mr. Trash Wheel sheet.

```{r}
path_mtw <- 
  "./dataset/Trash-Wheel-Collection-Totals-7-2020-2.xlsx"
mr_trashwheel <- 
  read_xlsx(
    path = path_mtw,
    sheet = "Mr. Trash Wheel") %>%                      #specify the sheet in the Excel file
  select(Dumpster:`Homes Powered*`) %>%                 #omit non-data entries
  janitor::clean_names() %>%
  drop_na(dumpster) %>%                                  #omit rows that do not include dumpster-specific data
  mutate(
    sports_balls = round(sports_balls, digits = 0)       #round the number of sports balls to the nearest integer
    ) %>%
  
```
  
* Step 2: Read and clean precipitation data for 2018 and 2019.

```{r}
# 2018 Precipitation
prec2018 <- 
  read_xlsx(
    path = path_mtw,
    sheet = "2018 Precipitation",                          
    skip = 1) %>%                       
  janitor::clean_names() %>%
  drop_na()                                              #omit rows without precipitation data

prec2018["year"] <- 2018                                 #add a variable for year

# 2019 Precipitation
prec2019 <- 
  read_xlsx(
    path = path_mtw,
    sheet = "2019 Precipitation",                          
    skip = 1) %>%                       
  janitor::clean_names() %>%
  drop_na()                                             #omit rows without precipitation data

prec2019["year"] <- 2019                                #add a variable for year
```
  
* Step 3: Combine precipitation datasets and convert month to a character variable. 
   
```{r}
perc_df <-
  rbind(prec2018,prec2019) %>%                          #combine precipitation datasets 
  mutate(
    month = month.name[month]
    ) %>%                                               #convert month to a character variable
  relocate(year,month)
```

* Step 4: Write a paragraph about these data.   
  
__This dataset contains three different trash wheel reports, which includes the time and amounts of various types of trash collected, and monthly precipitation report every year.__  
__For the Mr.Trash Wheel sheet, there are `r nrow(mr_trashwheel)` lines of observations after omitting rows that do not include dumpster-specific data. For yearly precipitation datasets, the total precipitation in 2018 is `r sum(prec2018$total)` inches. The median number of sports balls in a dumpster in 2019 was `r mr_trashwheel %>% filter(year == 2019) %>% pull(sports_balls) %>% median()`.__    
   
## Problem 2
   
* Step 1: Clean the data in pol-month.csv.   
   
```{r}
pol_month <- 
  read_csv("./dataset/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon,c("year","month","day")) %>%             #separate mon into integer year, month,day
  mutate(month = as.integer(month),
         year = as.integer(year),
         day = as.integer(day)) %>%
  mutate(month = month.name[month])                     #replace month number with month name 

pol_month["president"] <-
  ifelse(
    pol_month["prez_gop"] == "0", "dem","gop"           #president taking values gop and dem
    )     

pol_month <- 
  pol_month %>%
  select(-prez_dem,-prez_gop,-day)                     #remove prez_dem/gop and day 

```
    
* Step 2: Clean the data in snp.csv using a similar process to the above.   
   
```{r}
snp <- 
  read_csv("./dataset/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date,c("month","day","year")) %>%             #separate date into integer year, month,day
  mutate(month = as.integer(month),
         year = as.integer(year),
         day = as.integer(day),
         year = ifelse(
           year <= 21, year+2000,year+1900)) %>%         #change year format 
  mutate(month = month.name[month]) %>%                  #replace month number with month name 
  relocate("year","month","day","close") %>%             #year & month leading columns
  select(-day)                                           # remove day column 

```

* Step 3: Tidy the unemployment data
```{r}
unemp <- 
  read_csv("./dataset/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
  ) %>%
  rename( year = Year)

``` 
  
  * Step 4: Merge snp into pols and merge unemployment into result 
```{r}
snp_pols <-        
  left_join(pol_month,snp,by = c("year","month"))               #merge snp into pols

unemp_sp <-
  left_join(snp_pols,unemp,by = c("year","month"))             #merge unemployment 

```
  
* Step 5:Write a paragraph about these data     
    
__After cleaning the dataset, POLS_MONTH dataset contains 9 variables, which are YEAR, MONTH, GOV_GOP, SEN_GOP, REP_GOP, GOV_DEM, SEN_DEM, REP_DEM and PRESIDENT. There are `r nrow(pol_month)` lines of observations. The PRESIDENT variable is created based on PREZ_DEM and PREZ_GOP in the orginal dataset. People will be able to tell the number of national politicians who are democratic or republican at any given time. The range of years is from `r min(pol_month$year)` to `r max(pol_month$year)`.__
   
__After cleaning SNP dataset, it contains 3 variables, which are YEAR, MONTH and CLOSE. YEAR and MONTH will tell the date of observation while CLOSE will tell the closing values of the S&P stock index on the associated date. There are `r nrow(snp)` lines of observations. The range of years is from `r min(snp$year)` to `r max(snp$year)`.__     
   
__After tidying UNEMPLOYMENT dataset, it contains 3 variables, which are YEAR,MONTH,and UNEMPLOYMENT. Under UNEMPLOYMENT, people will be able to tell the percentage of unemployment in corresponding month of the associated year. There are `r nrow(unemp)`lines of observations and the range of years is from `r min(unemp$year)` to `r max(unemp$year)`.__  
    
## Problem 3   
    
* Step 1: Load and tidy the data   
    
```{r}
baby_name <- 
  read_csv("./dataset/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(
    childs_first_name = str_to_sentence(childs_first_name),                
    ethnicity = str_to_sentence(ethnicity),
    gender = str_to_sentence(gender),
    ethnicity = recode(ethnicity, 
                       `Asian and paci` = "Asian and pacific islander",
                       `Black non hisp` = "Black non hispanic",
                       `White non hisp` = "White non hispanic")) %>%
  distinct()

```
* Step 2: Produce a table - Olivia       
    
```{r}
olivia_df <-
  baby_name %>%
  filter(gender == "Female" &
           childs_first_name == "Olivia") %>%
  select(-gender,-count,-childs_first_name) %>%
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank) 

olivia_df <-
  olivia_df[,c("ethnicity","2011","2012","2013","2014","2015","2016")]

olivia_df  
```
* Step 3: Produce a similar table showing the most popular name among male children over time   
  
```{r}
male_df <-
  baby_name %>%
  filter(gender == "Male" &
           rank == "1") %>%
  select(year_of_birth,ethnicity,childs_first_name) %>%
  pivot_wider(
    names_from = year_of_birth,
    values_from = childs_first_name)

male_df <-
  male_df[,c("ethnicity","2011","2012","2013","2014","2015","2016")]

male_df
```
* Step 4: Produce a scatter plot   
  
```{r}
scatter_df <-
  baby_name %>%
  filter(year_of_birth == "2016" &
           gender == "Male" &
           ethnicity == "White non hispanic") %>%
  select(-year_of_birth,-gender,-ethnicity)

ggplot(scatter_df,
         aes(x = rank, y=count)) +
  geom_point() +
  ggtitle("Rank in popularity of a name vs. the number of childen with that name") +
  xlab("Rank in popularity") +
  ylab("Name count")

```


